---
title: "2015 Data Visualization"
author: "Radina Nikolic"
date: "Thursday, March 26, 2015"
output: html_document
---

## 1. Introduction

This document contains few Data Visualization Use Cases

## 2. Plane Crashes 

The first one was inspired by the latest plane crash in Europe. 

Data were extracted from database of plane acccidents hosted at: 

http://planecrashinfo.com.

Fro data aquisition R code posted by Phill Jette was used

https://github.com/philjette/CrashData/blob/master/PlaneCrashes.R

So here we started with preprocessed data.

```{r}

library(ggplot2)

setwd("C:/_RADINA/R/Planes")
crashData = read.csv("data/CompiledCrashData.csv", header=TRUE)
```


### Plane Crashes Data Visualization

```{r, echo=FALSE}

ggplot(crashData, aes(crashYear)) +
  geom_histogram(binwidth=1) +
  #fte_theme() +
  labs(x = "Year of Plane Crash", 
       y = "Number of Accidents") +
       scale_x_continuous(breaks=seq(1950,2015, by=5)) +
       scale_y_continuous(breaks=seq(0,110, by=5)) +
  ggtitle(expression(atop("Distribution of Plane Crashes (1950-2015)", 
                     atop(italic("Based on data from planecrashinfo.com"), ""))))  
```

### Plane Crashes Data Summary Statistics

```{r}
summary(crashData)
```

